---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

### Load Packages

```{r}

library(dplyr)
library(ggplot2)

```

### Load data

```{r echo = TRUE}
# Reading in CSV data using readr package and read_csv function

path <- file.path("activity", "activity.csv")

activity <- readr::read_csv(file = path)

```

<p>&nbsp;</p>
#### Examining file 

```{r echo=TRUE}
str(activity)

```

<p>&nbsp;</p>
#### Creating new file grouped by days across the two months and removing NA for this part of assignment.

```{r echo=TRUE}

Totalactivity <- activity %>% 
group_by(date) %>% 
  filter(!is.na(steps)) %>% 
  summarise(totalsteps=sum(steps))

```

## What is mean total number of steps taken per day?

<p>&nbsp;</p>
* Histogram showing total number of steps taken each day

```{r echo=TRUE}
Totalactivity %>% 
ggplot(mapping = aes(x = totalsteps)) +
  geom_histogram(fill = "blue")+
  labs(x ="Total numbers of steps in a day", y="Count")+
  ggtitle("History showing total number of steps taken each day")

```


<p>&nbsp;</p>
* Mean and median total number of steps taken per day across the two month period is calculated below.
```{r  echo=TRUE}
Totalactivity %>% 
  summarise(meansteps = mean(totalsteps), mediansteps = median(totalsteps))

```

The mean total number of steps per day was 10766 and the median total number of steps per day was 10765.

## What is the average daily activity pattern?

<p>&nbsp;</p>
* Time series plot of the 5-minute interval and the average number of steps taken.
```{r}
activity %>% 
  group_by(interval) %>% 
summarise(avgsteps = mean(steps, na.rm = TRUE)) %>% 
  ggplot(mapping = aes(x=interval, y=avgsteps))+
  geom_line(color="blue", size=0.7)+
  labs(x ="Interval", y="Average number of steps taken")+
  ggtitle("Plot showing Average Daily Activity Pattern")



```

<p>&nbsp;</p>
* The 5-minute interval, on average across all the days in the dataset, containing the maximum number of steps.

```{r}
activity %>% 
  group_by(interval) %>% 
summarise(avgsteps = mean(steps, na.rm = TRUE)) %>% 
            arrange(desc(avgsteps))
```
From the above, the 5-minute interval on average across all the days in the dataset, containing the maximum number of steps is the 835 interval.

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
